### 2024-08-24 README.qmd
Added Quarto markdown file for generating README.

---
title: "quicR"
format: gfm
editor: source
---

# Description
quicR is a package made for the analysis of real-time quaking-induced conversion (RT-QuIC) assays.

Much of the functionality is designed to be integrated with the Excel output files generated by the BMG analysis software, MARS.
For best results, in the MARS software select:

```{mermaid}
flowchart TB
excel(Excel Export)

subgraph A[Select the following export options]
  direction LR
  micro(Microplate View)
  table(Table View)
  trans(Transpose Table)
  test(Add test run \n information)
end

export(Export report to Excel)

excel --> A
micro --> table --> trans --> test
A --> export
```

Having both the microplate view and the table view helps with integrating plate layouts and the real-time data.

For many of the functions, it is important to have a table labelled "Sample IDs" on the microplate view sheet of the Excel file.